Topological Eulerian Synthesis of Slow Motion Periodic Videos

### 2018-02-05 Paper/paper.tex
Adding new shake results (may need some resolution/font tweaking
Labeling sections of pipeline as anaysis (a) and synthesis (b)

Figure~\ref{fig:ConceptFigure} shows the different parts of our method for a 1D periodic signal for illustration purposes. However, our technique applies to multivariate time series data as well, such as video. More specifically, the input to our method is a 3-channel video of $N$ frames, each frame of resolution $W \times H$. We treat each frame as a point in Euclidean space of dimension $\mathbb{R}^{3 \cdot W \cdot H}$, and denote the set of all videos by the sequence $\mb{X}(t) \in \mathbb{R}^{(3 \cdot W \cdot H)}$ for integer-valued $t \in [1,N]$. The steps of our approach outlined in Figure~\ref{fig:ConceptFigure}(a-c), however, do not rely on the {\em extrinsic} geometry of the set of points $\mb{X}(t)$, only its set of {\em pairwise distances}. Thus, we can project all of the points into a lower-dimensional space and apply the set of techniques to the projection, assuming that pairwise distances are preserved. Since $N \ll (3 \cdot W \cdot H)$, the frames can be projected into a $N-1$-dimensional space such that all distances are preserved. Denoting this projection by $\mb{P} \in \mathbb{R}^{(N-1) \times (3 \cdot W \cdot H)}$, our method for steps (a)-(c) operates on $\mb{Z}(t) = \mb{P} \mb{X}(t)$.  Furthermore, to mitigate the impact of drift from period to period, we use the third layer of a Gaussian pyramid on each frame in place of the raw frames before applying $\mb{P}$.

@@

The sliding window video embedding \cite{cao1998dynamics,traliehigh,tralie2017quasi} for $\mb{Z}$ is defined as
\begin{equation}
	SW_{d}[\mb{Z}(t)] = \left[ \begin{array}{c} \mb{Z}(t) \\ \mb{Z}(t + 1) \\ \vdots \\ \mb{Z}(t + (d-1))  \end{array} \right] \in \mathbb{R}^{(N-1)\cdot(d)},
\end{equation}
where $d$ is the dimension of the embedding. As shown by Takens~\cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(d-1)$ maximizes the roundness of the embeddings when $(d-1) = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D surrogate signal, and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate $D$.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{RobustnessTests.pdf}
\caption{Robustness tests.}
\label{fig:RobustnessTests}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{RobustnessTestsNoSW.pdf}
\caption{Robustness tests without sliding window.}
\label{fig:RobustnessTestsNoSW}
\end{figure}

We first experimentally quantify the accuracy of our circular coordinate inference, since we cannot hope to get a good slow motion template unless our circular coordinates are accurate.  We generate 3 different 600 frame synthetic periodic videos for which we know the ground truth circular coordinates, using software from \cite{jacobson2012fast}.  We vary the number of cycles that the videos undergo between 3, 5, 10, 15, 20, 25, 30, 40, and 50, and we also vary the ``shake'' of the video (width of a motion blur kernel) from 0, $20 \times 20$, $40 \times 40$, and $80 \times 80$, to assess the effect of drift.  We also compare different fractions $\kappa$ of mutual nearest neighbors to tune the threshold at which the unweighted Laplacian is built to TDA for aututuning the scale.  Figure~\ref{fig:RobustnessTests} shows the average angular error in degrees for our entire pipeline, and Figure~\ref{fig:RobustnessTests}.  The errors are overall quite low for moderate shake with a sliding window, though they increase as more shake is introduced.  Without a sliding window, the only video that performs reasonably well is the ``crowd'' videom, though the errors increase with shake much more rapidly than with the sliding window, supporting the ``time regularization'' properties of the sliding window.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{XTSliceFan.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of a fan with only 6 frames per period and the corresponding median consensus reordered video.}
\label{fig:XTSliceFan}
\end{figure}

>>>>

Figure~\ref{fig:ConceptFigure} shows the different parts of our method for a 1D periodic signal for illustration purposes. In general, the actual input to our method is a 3-channel video of $N$ frames, each frame of resolution $W \times H$. We treat each frame as a point in Euclidean space of dimension $\mathbb{R}^{3 \cdot W \cdot H}$, and denote the set of all videos by the sequence $\mb{X}(t) \in \mathbb{R}^{(3 \cdot W \cdot H)}$ for integer-valued $t \in [1,N]$. The steps of our approach outlined in Figure~\ref{fig:ConceptFigure}, which we split into two parts: an analysis part (a) and a synthesis part (b).  For the analysis part, we do not rely on the {\em extrinsic} geometry of the set of points $\mb{X}(t)$, only its set of {\em pairwise distances}. Thus, we can project all of the points into a lower-dimensional space and apply the set of techniques to the projection, assuming that pairwise distances are preserved. Since $N \ll (3 \cdot W \cdot H)$, the frames can be projected into a $N-1$-dimensional space such that all distances are preserved. Denoting this projection by $\mb{P} \in \mathbb{R}^{(N-1) \times (3 \cdot W \cdot H)}$, our method for the steps in (a) operates on $\mb{Z}(t) = \mb{P} \mb{X}(t)$.  We also use the third layer of a Gaussian pyramid on each frame in place of the raw frames before applying $\mb{P}$ to mitigate drift.

@@

The sliding window video embedding \cite{cao1998dynamics,traliehigh,tralie2017quasi} for $\mb{Z}$ is defined as
\begin{equation}
	SW_{d}[\mb{Z}(t)] = \left[ \begin{array}{c} \mb{Z}(t) \\ \mb{Z}(t + 1) \\ \vdots \\ \mb{Z}(t + (d-1))  \end{array} \right] \in \mathbb{R}^{(N-1)\cdot(d)},
\end{equation}
where $d$ is the dimension of the embedding. As shown by Takens~\cite{takens1981detecting}, a sliding window of dimension $2m+1$ of even a single generic observation function of a dynamical system of intrinsic dimension $m$ is sufficient to reconstruct a topological embedding of the underlying trajectory in the original state space.  For periodic signals, this state space is a torus, and $d$ should be twice the number of harmonics present to injectively reconstruct loops on that torus \cite{perea2015sliding}.  The same is true of videos in general \cite{tralie2017quasi}.  Furthermore, the sliding window length $(d-1)$ maximizes the roundness of the embeddings when $(d-1) = k T$ for some integer $k$, where $T$ is the fundamental period.  Since we will use topological data analysis (TDA) to find the scale of the loop, and since TDA measures roundness, it is important that we tune our window size to be an integer multiple of the period.  To do this, we perform 1D ISOMAP \cite{tenenbaum2000global} on the original frames of the video to generate a 1D surrogate signal, and we perform an autocorrelation-based fundamental frequency estimation \cite{Mcleod05asmarter} on this signal to estimate the period and, hence, the appropriate $d$.  This estimation often returns integer multiples of the fundamental period, which is fine for our purposes.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{RobustnessTestsShake.pdf}
\caption{Robustness tests.}
\label{fig:RobustnessTestsShake}
\end{figure}

We first experimentally quantify the accuracy of our circular coordinate inference, since we cannot hope to get a good slow motion template unless our circular coordinates are accurate.  We generate 3 different 600 frame synthetic periodic videos for which we know the ground truth circular coordinates, using software from \cite{jacobson2012fast}.  We vary the number of cycles that the videos undergo between 3, 5, 10, 15, 20, 25, 30, 40, and 50, and we also vary the ``shake'' of the video (width of a motion blur kernel) from $0 \times0 $ to $50 \times 50$ pixels, in increments of 10, to assess the effect of drift.  We also compare different fractions $\kappa$ of mutual nearest neighbors to tune the threshold at which the unweighted Laplacian is built to TDA for aututuning the scale.  Figure~\ref{fig:RobustnessTestsShake} shows the average angular error in degrees for our entire pipeline under different amounts of shake.  The errors are overall quite low for moderate shake with a sliding window, though they increase as more shake is introduced.  Without a sliding window, the only video that performs reasonably well is the ``crowd'' videom, though the errors increase with shake much more rapidly than with the sliding window, supporting the ``time regularization'' properties of the sliding window.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{XTSliceFan.pdf}
\caption{An XT slice of a line of pixels (magenta line, upper left) over time for an input video of a fan with only 6 frames per period and the corresponding reorderd template.}
\label{fig:XTSliceFan}
\end{figure}
